---
title: "Creately Assignment"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

# Prerequisites

## Import packages.

```{r}
library(tidyverse)
library(lubridate)
library(readxl)
library(ggplot2)
library(ggpubr)

setwd('C:/Personal/Data Analyst Job - Assingment/')
```
## Load Data.

```{r}
df <- read_excel("Dataset_DA_2022.xlsx")

glimpse(df)

```

## Data Cleaning

Both 'started_at' and 'cancelled_at' fields types are 'char'

```{r}
# Convert 'started_at' and 'cancelled_at' fields types to date
df <- mutate(df, 
             started_at = as_date(started_at, format = '%d/%m/%Y'),
             cancelled_at = as_date(cancelled_at, format = '%d/%m/%Y'))

glimpse(df)
```
Issue: It seems date field records are consisting with different formats.

```{r}
# Convert the excel file to a csv and load data from csv
df <- read_csv("Dataset_DA_2022.csv")

# Convert 'started_at' and 'cancelled_at' fields types to date 
df <- mutate(df, 
             started_at = as_date(started_at, format = '%d/%m/%Y'),
             cancelled_at = as_date(cancelled_at, format = '%d/%m/%Y'))

glimpse(df)
#tail(df, 100)
```
Check for null values

```{r}
dim(df)

df %>% filter(is.na(CUID))

```

There are 1,020,251 null records from field CUID

```{r}
# Removing null records
df <-  df %>% filter(!is.na(CUID))


# Verify null record removal
df %>% filter(is.na(CUID)) %>% count()
df %>% filter(is.na(subscription)) %>% count()
df %>% filter(is.na(started_at)) %>% count()
df %>% filter(is.na(cancelled_at)) %>% count()

dim(df)
```
Null values removed
8368 records with data.

# Feature Engineering

## Check subscription types

```{r}
df %>% select(subscription) %>% unique()
```

There are 6 subscription types in the data sets.

```{r}
df

summary(df)
```
## Introduce new features

```{r}
# Introduce new feilds
# usage_days (cancelled_at - started_at)
# started_month
df <- mutate(df, 
             usage_days = difftime(cancelled_at, started_at, units = "days"),
             started_month = month(ymd(started_at), 
                                   label = FALSE, 
                                   abbr = FALSE),
             started_month_nm = month(ymd(started_at), 
                                      label = TRUE, 
                                      abbr = FALSE))

df$usage_days <- as.numeric(df$usage_days, units="days")
df$usage_months <- as.numeric(ceiling(df$usage_days / 30), units = "month")

summary(df)
```
New fields: 
usage_days - usage period in days
usage_months - usage period in months
started_month - subscribed month
started_month_nm - subscribed month name

```{r}
summary(df)
```
# Detailed Analysis

## 1) Sales

```{r}
subscriptions <- df %>% group_by(subscription) %>% 
  count(name = 'No_of_subscriptions')
subscriptions

total_subs <- as.numeric(count(df))
 
subscriptions = subscriptions %>% mutate(
  subscribed_percentage = No_of_subscriptions / total_subs*100)

ggplot(subscriptions, aes(y = reorder(subscription, No_of_subscriptions), 
                        x = subscribed_percentage))+ 
  geom_bar(stat = "summary")+
  scale_x_continuous(position = "top") +
  geom_text(aes(label = paste0(round(subscribed_percentage,1), " %")), colour = "white", hjust = 1.5)+
  scale_fill_manual(name = "Reviews", values=c("#F7C815","grey50")) +
  labs(title = "Subscriptions",
       x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
    plot.caption = element_text(face = 'italic'),
    panel.grid.major = element_line('white', size = 0.5),
    panel.grid.minor = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.ontop = FALSE
  )

#ggsave("1_overall_subcriptions.jpg", width = 10, height = 8, units = "cm")
```
It is a clear highlight that, 'Plus-monthly' is the most commonly subscribed package.

```{r}
df %>% group_by(subscription) %>% count()
```

This shows that only 'Plus-monthly', 'Growth-monthly' and 'Plus-yearly' subscription types have a minimal of 100 subscriptions.
Other subscription types not even exceed 10 subscriptions.

### Overall subscriptions by month

```{r}
monthly_subscriptions3 <- df %>% 
  group_by(started_month_nm) %>% 
  count() %>% 
  arrange(started_month_nm)

total_subs <- as.numeric(count(df))
 
monthly_subscriptions3 = monthly_subscriptions3 %>% mutate(
  subscribed_percentage = n / total_subs*100)


monthly_subscriptions3

#sum(monthly_subscriptions3$subscribed_percentage)

```

```{r}
ggplot(monthly_subscriptions3, aes(y = subscribed_percentage, 
                      x = started_month_nm))+ 
  geom_bar(stat = "summary")+
  labs(title = "Subscriptions by Month",
       x = NULL,
       y = 'Percentage %') +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
    plot.caption = element_text(face = 'italic'),
    panel.grid.major = element_line('white', size = 0.5),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.ontop = TRUE
  )

#ggsave("2_subscriptions_by_month.jpg", width = 16, height = 12, units = "cm")
```
Overall it indicate a good subscription sales from January to April. Then there is a considerable drop of sales.

### Subscriptions by Subscription Types

```{r}
monthly_subscriptions <- df %>% filter((subscription == 'Growth-monthly')|
                                       (subscription == 'Plus-monthly')|
                                       (subscription == 'Plus-yearly')) %>%
  group_by(started_month_nm, subscription) %>%
  count(subscription, name = "n") %>%
  pivot_wider(names_from = subscription, values_from = n, values_fill = 0) %>%
  ungroup()

monthly_subscriptions <-data.frame(monthly_subscriptions, row.names = 1)

growth_m_sum <- sum(monthly_subscriptions$Growth.monthly)
plus_m_sum <- sum(monthly_subscriptions$Plus.monthly)
plus_y_sum <- sum(monthly_subscriptions$Plus.yearly)

monthly_subscriptions$Growth.monthly <- monthly_subscriptions$Growth.monthly / growth_m_sum * 100
monthly_subscriptions$Plus.monthly <- monthly_subscriptions$Plus.monthly / plus_m_sum * 100
monthly_subscriptions$Plus.yearly <- monthly_subscriptions$Plus.yearly / plus_y_sum * 100


head(monthly_subscriptions)

ggballoonplot(monthly_subscriptions, fill = "value", size.range = c(1, 8)) +
  scale_fill_viridis_c(option = "C")+
  guides(size = FALSE) + 
  labs(title = "Percentages: Subcriptions (Min 100 subscriptions)",
       fill = 'Percentage (%)',
       x = NULL,
       y = NULL)
```
This shows the percentages of sales per each subscription type (top 3).
But this is not the best graphical representation.

Another approach.

```{r}
monthly_subscriptions2 <- df %>% 
  filter((subscription == 'Growth-monthly')| (subscription == 'Plus-monthly')| (subscription == 'Plus-yearly')) %>%
  group_by(subscription, started_month_nm) %>% 
  count() %>% 
  arrange(subscription, started_month_nm)
 
monthly_subscriptions2 = monthly_subscriptions2 %>% mutate(
  subscribed_percentage = case_when(subscription == "Growth-monthly"~ n/growth_m_sum*100,
                                    subscription == "Plus-monthly"~ n/ plus_m_sum * 100,
                                    subscription == "Plus-yearly"~ n/plus_y_sum * 100)
 )


monthly_subscriptions2

```

```{r}
ggplot(monthly_subscriptions2, aes(y = subscribed_percentage, 
                      x = started_month_nm, fill = subscription))+ 
  geom_bar(stat = "summary")+
  facet_wrap(~ subscription, ncol= 1)+
  labs(title = "Subscriptions by Month",
       x = NULL,
       y = '%') +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
    plot.caption = element_text(face = 'italic'),
    panel.grid.major = element_line('white', size = 0.5),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.ontop = TRUE
  )

#ggsave("3_subscription_sales_by_subscription_type.jpg", width = 16, height = 16, units = "cm")
```

## 2) Subscription cancellations

### Average subscription usage period of each subscription type.

```{r}
usage <- df %>% group_by(subscription) %>% summarise(average_usage_days = mean(usage_days))
#usage

ggplot(usage, aes(x = reorder(subscription, -average_usage_days), y = average_usage_days)) +
  geom_bar(stat = "summary", fun = "mean") +
  labs(title = "Average Usage Days",
       x = NULL,
       y = "Days")+
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
    plot.caption = element_text(face = 'italic'),
    panel.grid.major = element_line('white', size = 0.5),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_blank(),
    panel.ontop = TRUE
  )

#ggsave("4_", width = 16, height = 16, units = "cm")
```
This indicate that only 'Plus-monthly' and 'Growth-monthly' subscribers use the service comparatively longer than other subscription types.
But still the average usage duration is less than 50 days.

### Spread of subscription usage period by each subscription type.

```{r}
ggplot(df) + 
  geom_boxplot(aes(x = usage_days, y = subscription, fill = subscription))+
  #scale_x_continuous(limits = c(0, 5000)) + 
  labs(title = "Usage Distribution",
       x = "Usage Days",
       y = NULL) +
  theme_minimal() +
  theme(legend.position="none")

#ggsave("5_subscription_usage_distribution.jpg", width = 16, height = 12, units = "cm")
```
This indicate that 'Plus-monthly', 'Growth-monthly' and 'Plus-yearly' subscription types shows wider usage spread.
It would be easy to understand view subscription cancellation percentages by usage months for each subscription type.

### Subscription cancellation percenages by period of months

Filter subscription types with minimum 100 subscribers.
```{r}
# Alter data
min_usage <- df %>% filter((subscription == 'Growth-monthly')|
                                       (subscription == 'Plus-monthly')|
                                       (subscription == 'Plus-yearly')) %>%
  group_by(subscription, usage_months) %>% 
  count() %>% 
  arrange(subscription, usage_months)
 
min_usage = min_usage %>% mutate(
  cancel_percentage = case_when(subscription == "Growth-monthly"~ n/growth_m_sum*100,
                                subscription == "Plus-monthly"~ n/ plus_m_sum * 100,
                                subscription == "Plus-yearly"~ n/plus_y_sum * 100)
 )

min_usage = min_usage %>% mutate(
  cancellation_type = case_when(usage_months == 0~ "Same Day",
                                usage_months <= 1~ "Within 1 Month",
                                usage_months <= 2~ "Within 2 Months",
                                usage_months <= 3~ "Within 3 Months",
                                usage_months <= 4~ "Within 4 Months",
                                usage_months <= 5~ "Within 5 Months",
                                usage_months <= 6~ "Within 6 Months",
                                usage_months <= 7~ "Within 7 Months")
 )

# Verify percentages

test <- min_usage %>% filter(subscription == "Growth-monthly") 
test
sum(test$cancel_percentage)

test <- min_usage %>% filter(subscription == "Plus-monthly") 
test
sum(test$cancel_percentage)

test <- min_usage %>% filter(subscription == "Plus-yearly") 
test
sum(test$cancel_percentage)
```

```{r}
ggplot(min_usage, aes(y = cancel_percentage, 
                      x = cancellation_type))+ 
  geom_bar(stat = "summary")+
  facet_wrap(~ subscription, ncol= 1)+
  labs(title = "Subscription Cancellations",
       x = NULL,
       y = 'Percentage') +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
    axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)
  )
```
This shows subscription cancellation percentages for each subscription type by usage period.
But this is not the perfect view to present inside.

Note: There are considerable percentage of same day cancellations for "Plus-yearly" subscription type.

I want to show how sooner the business looses their subscribers.
It would be more useful to use cumulative subscription cancellation percentage rather than individual percentages by usage period.

### Cumulative subscription cancellation percentages by period of months

Overall subscription cancellations

```{r}
min_usage2 <- df %>% filter((subscription == 'Growth-monthly')|
                            (subscription == 'Plus-monthly')|
                            (subscription == 'Plus-yearly')) %>%
  group_by(subscription, usage_months) %>%
  summarise(proportion = n()) %>%
  mutate(Perc = cumsum(100*proportion/sum(proportion))) %>%
  select(-proportion)

ggplot(min_usage2, aes(y = Perc, 
                      x = usage_months))+ 
  geom_point(stat = "summary") + 
  geom_line(stat = "summary") + 
  geom_vline(xintercept = 4, size = 4, alpha = 0.15) + 
  geom_vline(xintercept = 0, size = 4, alpha = 0.15) + 
  labs(title = "Subscription Cancellations",
       x = 'Usage Months',
       y = 'Cancellations %') +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
  )

ggsave("6_overall_subscription_cancellations.jpg", width = 16, height = 12, units = "cm")
```
This shows that,
Overall the business looses almost 100% of the subscribers after 4 months
There is about 6-8% same day cancellation subscribers.

### Cumulative subscription cancellation percentages by period of months per each subscription type


```{r}
min_usage3 <- df %>% filter((subscription == 'Growth-monthly')|
                            (subscription == 'Plus-monthly')|
                            (subscription == 'Plus-yearly')) %>%
  group_by(subscription, usage_months) %>%
  summarise(proportion = n()) %>%
  mutate(Perc = cumsum(100*proportion/sum(proportion))) %>%
  select(-proportion)


min_usage3
```
```{r}
ggplot(min_usage3, aes(y = Perc, 
                      x = usage_months))+ 
  geom_point(stat = "summary")+
  geom_line(stat = "summary")+
  facet_wrap(~ subscription, ncol= 1)+
  labs(title = "Subscription Cancellations",
       x = NULL,
       y = 'Percentage') +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
  )
```
This shows the cumulative cancellation percentages by each subscription type. 
It is noticeable that 'Plus-yearly' package losses its subscribers more sooner than other two types.

But still this does not show gives the seriousness of the issue.
It would be better to look at things in a different angle.
It will give a strong message if I show how long it will take to loose a certain percentage of subscribers.


### Time duration to loose 90% of subscribers per each subscription type

```{r}
loss_3_months <- min_usage2 %>% filter(usage_months <= 3)  %>% slice_max(Perc)
loss_3_months
```
```{r}
months_90_cancellations <- min_usage2 %>% filter(Perc >= 89) %>% slice(1) # 89% rounded up to 90%
months_90_cancellations
```


```{r}
ggplot(months_90_cancellations, aes(y = subscription, 
                      x = usage_months))+ 
  geom_bar(stat = "summary")+
  #geom_text(aes(label = usage_months), colour = "white", hjust = 3)+ #paste0(seq(0, 0.6, by = 0.1), "%")
  geom_text(aes(label = paste0(usage_months, " month")), colour = "white", hjust = 1.5)+ 
  labs(title = "Duration to Lose 90% Subscribers",
       #x = 'Months',
       x = NULL,
       y = NULL) +
  theme_minimal() +
  theme(
    strip.text = element_text(face = 'bold', hjust = 0),
    plot.caption = element_text(face = 'italic'),
    panel.grid.major = element_line('white', size = 0.5),
    panel.grid.minor = element_blank(),
    axis.text.x=element_blank(),
    panel.grid.major.x = element_blank(),
    panel.ontop = FALSE
  )

#ggsave("7_duration_90_percent_lose_subscribers.jpg", width = 12, height = 8, units = "cm")
```
This plots is more meaningful.

'Plus-yearly' subscription type looses 90% of their subscribers just within 1 month.

Both 'Plus-monthly' and 'Growth-monthly' subscription packages looses 90% of their subscribers within 3 months.

